---
title: "Exercises"
author: "Jacob Pammer, Chandler Wann, Narain Mandyam, Rudraksh Garg"
date: "8/1/2021"
output: pdf_document
---

```{r Problem 1, echo=FALSE, include = FALSE}

df = read.csv(file = "data/greenbuildings.csv")

table(df$green_rating)
hist(df$leasing_rate)

df = df[df$leasing_rate > 10,]
hist(df$leasing_rate)

Monthly_Revenue = (df$Rent * df$size * (df$leasing_rate/100))
df$Monthly_Revenue_in_Mil = Monthly_Revenue/1000000

drops <- c("LEED","Energystar")
df = df[ , !(names(df) %in% drops)]

df$Rent_Diff = df$Rent - df$cluster_rent


df_green = df[df$green_rating == 1,]
df_nongreen = df[df$green_rating == 0,]

median(df_green$Rent)
median(df_nongreen$Rent)

median(df_green$Rent) - median(df_nongreen$Rent)

IQR(df_green$Rent)
IQR(df_nongreen$Rent)

par(mfrow=c(1,2))

d0 <- seq(0, 200, by=10)
d1 <- seq(0, 200, by=10)
lmts <- range(d0,d1)



boxplot(df_green$Rent, col='green', ylim= lmts)

boxplot(df_nongreen$Rent, ylim=lmts)



wilcox.test(df_green$Rent, y = df_nongreen$Rent, alternative = c("greater"))





```
```{r Problem 1, echo=FALSE}
plot(df$green_rating, df$age)



```

ANSWER HERE::
Do we aggree with conclusions?
Certain portions of the analysis were done correctly, such as the use of medians is a good idea to avoid outliers.
Additionally, the according to the wilcoxon test we found that the medians of green vs non-green buildings were statistically different. So,
the green buildings could potentially be a strong indicator for rent price in the aggregate.
However, the analyst did not account for certain cities being more green then others. Certain clusters will have higher rents regaurdless of green. Therefore,
failing to take this into account could be undercutting your additionaly revenue. On top of that, in your payback period the analyst failed to consider the time value of money. 
Instead of creating a straight line payback, the analyst should use a discounted payback period or NPV. This would determine the payoff period and addional cashflows more accurately. 
It would be nice for comparison purposes if the clusters were correlated to certain cities. 
(bar plot could comparing npv vs non npv adjustment)
Then find a good scatter plot comparing stuff and discuss this as confounding
Even though there are many flaws in this study, choosing a green building could be a viable option, but further
analysis that address our confounding variables would be needed to be sure.




```{r Problem 1, echo = False}
#names(df)
#median(df$leasing_rate)
plot(df$green_rating, df$cluster_rent) 


```
